Local Adaptive Optimization of Time step

### 2013-10-22 report/contract.tex
Litterature study handin and papers
Cleanup of code and report

\section{Learning Goals}
Course specific learning goals:
\begin{enumerate}
	\item Formulate an operational project plan.
	\item Search the relevant literature and write a literature review setting own work in perspective.
	\item Be able to formulate what plagiarism is, and demonstrate proper citation and reference style.
	\item Solve a selected problem of fair Computational and Mathematical Modelling / eScience content and difficulty.
	\item Produce a thorough experiment plan that clearly demonstrates the quality of and highlights boundaries for the solution. 
	\item Produce a scientific text of fair quality both textually and scientifically.
	\item The student must be able to make an oral presentation of own work.
\end{enumerate}
Project specific learning goals:
\begin{enumerate}
    \item Implement local adaptive optimization of timesteps (LAOT)
	\item Evaluate LAOT to minimize the risk of unwanted deformations in grids.
    \item Discuss how LAOT is performed in practice using existing 2nd order optimization methods.
    \item Study of convergence on implemented methods.
\end{enumerate}

@@

\subsection{Problem Statement}
Study the application of locally adaptive timestepping in simulation of hyperelastic materials.

\subsection{Motivation}
When running a simulation of large hyperelastic materials we wish to make
large enough steps for the simulation to be performed within a reasonable
timeframe, but we still want the simulation to be accurate and to avoid
the mesh from curling up or exploding due to too large determinants of the
moving mesh vertices. In some hyperelastic materials most of the
discretized points are able to be simulated using larger timesteps whereas
smaller areas might need smaller timesteps. These situations are the ones
we wish to accomodate in this paper.

@@

\subsection{Risk assessment}
Low to none. The technique has been performed before in various shapes and it
should therefore be do-able.

>>>>

\section{Learning Goals}
Course specific learning goals:
\begin{enumerate}
	\item Formulate an operational project plan.
	\item Search the relevant literature and write a literature review setting own work in perspective.
	\item Be able to formulate what plagiarism is, and demonstrate proper citation and reference style.
	\item Solve a selected problem of fair Computational and Mathematical Modelling / eScience content and difficulty.
	\item Produce a thorough experiment plan that clearly demonstrates the quality of and highlights boundaries for the solution. 
	\item Produce a scientific text of fair quality both textually and scientifically.
	\item The student must be able to make an oral presentation of own work.
\end{enumerate}
Project specific learning goals:
\begin{enumerate}
    \item Implement local adaptive optimization of timesteps (LAOT)
	\item Evaluate LAOT to minimize the risk of unwanted deformations in grids
        and minimize the computational cost.
    \item Discuss how LAOT is performed in practice using existing 2nd order optimization methods.
\end{enumerate}

@@

\subsection{Problem Statement}
Study the application of locally adaptive timestepping in 1 dimensional
simulation of a mass spring system.

\subsection{Motivation}
When running a simulation of large hyperelastic materials we wish to make
large enough steps for the simulation to be performed within a reasonable
timeframe, but we still want the simulation to be accurate and to avoid
the mesh from curling up or exploding due to too large determinants of the
moving mesh vertices. In some hyperelastic materials most of the
discretized points are able to be simulated using larger timesteps whereas
smaller areas might need smaller timesteps. These situations are the ones
we wish to accomodate in this paper by gaining insight into the area of LAOT in
a simple 1 dimensional simulation (mass spring system).

@@

\subsection{Risk assessment}
Medium. The technique of LAOT has been performed before in various shapes but
not the same as our proposed solution.

### 2013-10-22 report/report.tex
Cleanup of code and report

\section{Introduction}
Simulations of large (order of millions of vertices) hyper elastic materials
are very time consuming. When working with materials of varying density
of vertices, the highest densities often make the most strict bounds on
size of the time step of the simulation in order to keep the materials (and
grids) stable and the discretization correct. In a ``standard'' simulation
with a global time step size we need to compute possibly unnecessarily many
computations on the coarser parts of the material caused by this bound.
When furthermore having materials with a large amount of somewhat equally
distributed vertices and smaller patches of detailed areas with higher density
of vertices, it would be an advantage to be able to perform large time steps
for the majority of the material and smaller steps for the local patches when
vertices cause instability. We call this concept local adaptive optimization
of time step, which we will study further in this report.

@@

\subsection{Initial though}
Normal timestepping with switch detection and refinement

\subsection{Stepping with interpolation and lockout of re-calculations}

@@

>>>>

\section{Introduction}
simulations of large (order of millions of vertices) hyper elastic materials
are very time consuming. when working with materials of varying density
of vertices, the highest densities often make the most strict bounds on
size of the time step of the simulation in order to keep the materials (and
grids) stable and the discretization correct. in a ``standard'' simulation
with a global time step size we need to compute possibly unnecessarily many
computations on the coarser parts of the material caused by this bound.
when furthermore having materials with a large amount of somewhat equally
distributed vertices and smaller patches of detailed areas with higher density
of vertices, it would be an advantage to be able to perform large time steps
for the majority of the material and smaller steps for the local patches when
vertices cause instability. we call this concept local adaptive optimization
of time step, which we will study further in this report.

@@

\subsection{Switch}

\subsection{Spatial LAOT}

\subsection{Inverse}

@@

Adaptive timestepping based on highest acceleration/smallest grid
